NumPy---Array Programming in Python

### 2020-02-20 body.tex
Minor edits

% KJM: I moved this here to get the placement right
\begin{figure*}[h]
  \centering
  \includegraphics[width=\textwidth]{static/sketches/concepts}
  \caption{\textbf{The NumPy array incorporates several fundamental array concepts.}
    \textbf{a,} The NumPy array data structure and its associated metadata fields.
    \textbf{b,} Indexing an array with slices and steps. These
  operations return a \emph{view} of the original data.
    \textbf{c,} Indexing an array with masks, scalar coordinates, or
  other arrays, so that it returns a copy of the original data. In the
  bottom example, an array is indexed with other arrays; this broadcasts the indexing arguments before performing the lookup.
    \textbf{d,} Vectorization efficiently applies operations to groups
  of elements.
    \textbf{e,} Broadcasting in the multiplication of two-dimensional arrays.
    \textbf{f,} Reduction operations act along one or more axes. In this
    example, an array is be summed along various axes to produce a vector, or along two axes consecutively to
    produce a scalar.
    \textbf{g,} Example NumPy code, illustrating some of these concepts.
   }
  \label{fig:array-concepts}
\end{figure*}

@@

NumPy combines the expressive power of \emph{array programming},
the performance of C, and
the readability, usability, and versatility of Python in a mature,
well-tested, well-documented, and community-developed library.
Many fast algorithms are implemented by libraries in the
scientific Python ecosystem.
Where extreme optimization is warranted, compiled languages such as
Cython \cite{behnel2011cython}, Numba \cite{Lam:2015:NLP:2833157.2833162},
and Pythran \cite{guelton2015pythran}, that
extend Python and transparently accelerate bottlenecks, can be
used.
Because of NumPy's simple memory model, it is easy to write low-level, hand-optimized code, usually in C
or Fortran, to manipulate NumPy arrays and pass them back to
Python.
Furthermore, using array protocols, it is possible to utilize the full
spectrum of specialized hardware acceleration with minimal changes to
existing code.

@@

Over the next decade, we will face several new challenges.  New devices will be
developed, and existing specialized hardware will evolve, to meet diminishing
returns on Moore's law.  There will be more, and a wider variety of, data
science practitioners, a significant proportion of whom will be using NumPy.
The scale of scientific data gathering will continue to expand, with the
adoption of devices and instruments such as light sheet microscopes and the
Large Synoptic Survey Telescope (LSST) \cite{jenness2018lsst}.  New generation
languages, interpreters, and compilers, such as Rust
\cite{10.1145/2692956.2663188}, Julia \cite{Julia-2017}, and LLVM
\cite{LLVM:CGO04}, will invent and determine the viability of new concepts and
data structures.

>>>>

% KJM: I moved this here to get the placement right
\begin{figure*}[h]
  \centering
  \includegraphics[width=\textwidth]{static/sketches/concepts}
  \caption{\textbf{The NumPy array incorporates several fundamental array concepts.}
    \textbf{a,} The NumPy array data structure and its associated metadata fields.
    \textbf{b,} Indexing an array with slices and steps. These
  operations return a \emph{view} of the original data.
    \textbf{c,} Indexing an array with masks, scalar coordinates, or
  other arrays, so that it returns a copy of the original data. In the
  bottom example, an array is indexed with other arrays; this broadcasts the indexing arguments before performing the lookup.
    \textbf{d,} Vectorization efficiently applies operations to groups
  of elements.
    \textbf{e,} Broadcasting in the multiplication of two-dimensional arrays.
    \textbf{f,} Reduction operations act along one or more axes. In this
    example, an array is summed along select axes to produce a vector, or along two axes consecutively to
    produce a scalar.
    \textbf{g,} Example NumPy code, illustrating some of these concepts.
   }
  \label{fig:array-concepts}
\end{figure*}

@@

NumPy combines the expressive power of \emph{array programming},
the performance of C, and
the readability, usability, and versatility of Python in a mature,
well-tested, well-documented, and community-developed library.
Libraries in the scientific Python ecosystem provide fast implementations of most important algorithms.
Where extreme optimization is warranted, compiled languages such as
Cython \cite{behnel2011cython}, Numba \cite{Lam:2015:NLP:2833157.2833162},
and Pythran \cite{guelton2015pythran}, that
extend Python and transparently accelerate bottlenecks, can be
used.
Because of NumPy's simple memory model, it is easy to write low-level, hand-optimized code, usually in C
or Fortran, to manipulate NumPy arrays and pass them back to
Python.
Furthermore, using array protocols, it is possible to utilize the full
spectrum of specialized hardware acceleration with minimal changes to
existing code.

@@

Over the next decade, we will face several challenges.  New devices will be
developed, and existing specialized hardware will evolve, to meet diminishing
returns on Moore's law.  There will be more, and a wider variety of, data
science practitioners, a significant proportion of whom will be using NumPy.
The scale of scientific data gathering will continue to expand, with the
adoption of devices and instruments such as light sheet microscopes and the
Large Synoptic Survey Telescope (LSST) \cite{jenness2018lsst}.  New generation
languages, interpreters, and compilers, such as Rust
\cite{10.1145/2692956.2663188}, Julia \cite{Julia-2017}, and LLVM
\cite{LLVM:CGO04}, will invent and determine the viability of new concepts and
data structures.
